Corotating Dust Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-06-29 paper/ms.tex
update ms

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep{2010exop.book...55W}.

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show quasiperiodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle; sharp peaks
are common, but .  Most CQVs are pre-main-sequence M-dwarfs without
near-infrared excesses, with ages of $\approx$5-150 million years
(Myr), and rotation periods of at most two days; they are observed to
be $\approx$1-3\% of M-dwarfs younger than 100 Myr
\citep{2016AJ....152..114R,2022AJ....163..144G}.  The dips can be
chromatic, with a reddening law plausibly consistent with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

@@

In our view, the arguments seem best established for the dust clump
hypothesis \citep{2023MNRAS.518.4734S}.  This is primarily because the
prominence scenario lacks a mechanism for generating broadband
opacity, and the screen scenario seems inconsistent with the degree of
observed periodicity, and the lack of infrared excess.  However,
unambiguous evidence, for instance a spectroscopic detection of
silicates during a CQV ``dip'', has yet to be presented to confirm
which scenario is correct.

@@

We visually assessed whether the objects found using the Fourier
(Section~\ref{subsec:fourier}) and dip-counting
(Section~\ref{subsec:counting}) techniques were consistent with
expectations for CQVs by assembling the data shown in
Appendix~\ref{app:vetting}.

>>>>

All pre-main-sequence stars vary in optical brightness, and the origin
of such variability is, in most cases, understood.  Well-explored
sources of optical variability include inhomogeneities on stellar
surfaces such as starspots and faculae \citep{2021isma.book.....B},
occultations by gas-rich circumstellar disks
\citep{2017MNRAS.470..202B}, and, in geometrically favorable
circumstances, eclipses by stars and planets
\citep{2010exop.book...55W}.  More exotic forms of optical variability
relevant to this work include transiting exocomets (e.g. $\beta$~Pic;
\citealt{2019A&A...625L..13Z}) and disintegrating rocky bodies around
both M-dwarfs (e.g. K2-22; \citealt{2014ApJ...784...40R}) and white
dwarfs (e.g. WD~1145; \citealt{2015Natur.526..546V}).

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show nearly periodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle.  Most CQVs
are pre-main-sequence M-dwarfs without near-infrared excesses, with
ages of $\approx$5-150 million years (Myr), and rotation periods of at
most two days; they are observed to be $\approx$1-3\% of M-dwarfs
younger than 100 Myr \citep{2016AJ....152..114R,2022AJ....163..144G}.
The dips can be chromatic, with a reddening law plausibly consistent
with dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

@@

In our view, the arguments seem best established for the dust clump
hypothesis \citep{2023MNRAS.518.4734S}.  This is primarily because the
prominence scenario lacks a mechanism for generating broadband
opacity, and the screen scenario is inconsistent with the degree of
observed periodicity, the lack of infrared excess, and the observed
lifetime of the effect, which is an order of magnitude longer than the
$\approx$$10^7$ years timescale typically quoted for the primordial
disk to disperse (CITE).  However, unambiguous evidence, for instance
a spectroscopic detection of silicates during a ``dip'', has yet to be
acquired.

@@

We visually assessed whether the objects found using the Fourier
(Section~\ref{subsec:fourier}) and dip-counting
(Section~\ref{subsec:counting}) techniques were consistent with
expectations for CQVs by assembling the data shown in
Appendix~\ref{app:vetting}.
We labelled a star as a ``good'' CQV if at least one TESS sector
showed what we viewed as the unambiguous signatures of the class (short
period; at least three dips or else otherwise oddly-shaped dips;
relative stability over a timescale of 30\,days).
We also noted stars that we thought could be CQVs, but that were more
ambiguous with a ``maybe'' flag.
